---
title: "ECON233 Project"
author: "Tho Do, Yang Zhiyao, Soh Zhe Hong"
date: "April 1, 2019"
output:
  html_document: default
  word_document: default
---
We investigate the components of Singapore Domestic Export time series from Jan 1976 to Dec 2017. Then, we attempt to forecast it for the period of Jan 2018-Dec2018.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r echo=FALSE, message=FALSE}
# This note uses the following libraries
# find the path for Dropbox across machines
library(jsonlite)
file_name<-list.files(paste(Sys.getenv(x = "APPDATA"),"Dropbox", sep="/"), pattern = "*.json", full.names = T)
if (length(file_name)==0){
  file_name<-list.files(paste(Sys.getenv(x = "LOCALAPPDATA"),"Dropbox", sep="/"), pattern = "*.json", full.names = T)}

dropbox_path<-fromJSON(txt=file_name)$personal
dropbox_path<-dropbox_path$path

# Import necessary packages
library(tidyverse)
library(forecast)
library(ggfortify)
library(gridExtra)
library(grid)
library(dynlm)
library(xts)
library(lubridate)
library(urca)
source(paste(dropbox_path,"\\ECON233\\","util.R",sep="")) # a R script containing all util functions from Prof's notes
```

## 1. Data Transformation

Input data and transform data timeseries object.
```{r echo=FALSE, message=FALSE}
# read the data
dx_data <- read_csv(paste(dropbox_path,"\\ECON233\\","dx_2017.csv",sep=""), col_names = FALSE)
colnames(dx_data) <- c("Date", "DX")
#dx_data$Date <- as.vector(dx_data$Date)

# convert to xts
#dx_xts <- xts(select(dx_data, -Date), order.by = as.yearmon(dx_data$Date, "%b-%y"))
start_date <- as.yearmon(dx_data$Date[1],"%b-%y")
end_date <- as.yearmon(dx_data$Date[nrow(dx_data)],"%b-%y")
dx_ts <- ts(dx_data$DX, start = c(year(start_date), month(start_date)),
                              end = c(year(end_date), month(end_date)),
                              frequency = 12)
```

```{r message = FALSE}
# transform to log 
p1 <- autoplot(dx_ts) + xlab("Domestic Export Singapore") + theme(aspect.ratio=1/2)
y <- ts(log(dx_ts), start = c(year(start_date), month(start_date)),
        end = c(year(end_date), month(end_date)),
        frequency = 12)
p2 <- autoplot(y) + xlab("Domestic Export Singapore") + theme(aspect.ratio=1/2)
grid.arrange(p1,p2, ncol = 2)
correl(y)
```

Run the ADF and KPSS test to check stationary
```{r message = FALSE}
y %>%  ur.df(type="trend", selectlags="AIC") %>% summary()
y %>% ur.kpss(type="tau", lags="short") %>% summary()
```

The series is not stationary from both tests. Hence we proceed to first difference it. We run similar tests on differenced series to ensure that it is now stationary.

```{r message = FALSE}
diff(y) %>% ur.df(type = "drift", selectlags = "AIC") %>% summary()
diff(y) %>% ur.kpss(type ="mu", lags = "short") %>% summary()
correl(diff(y))
```


PACF, we see 2 significant spikes at lag 1 and 2. Hence, we will fit an AR(2) component in our model.

```{r echo=FALSE, message = FALSE}
# Define trend and seasonals
trendvar <- seq_along(y)
seasonalvars <- seasonaldummy(y)
trendseasvars = cbind(trendvar, seasonalvars)
```

We investigate seasonality to see whether it exists

```{r message = FALSE}
ggseasonplot(diff(y)) + theme(aspect.ratio = 1/2)
ggsubseriesplot(diff(y)) + ggtitle("Monthly mean of DX") + theme(aspect.ratio = 1/2) # a seasonality component on March
```

There are seasonality at March every year. This suggests us to try either SARIMA or ARIMA with seasonal dummies.

## 2. Model Selection

We run different models and pick the one with smallest BIC.

```{r message = FALSE}
fit1 <- Arima(y, order=c(2,1,0),
              include.constant = T,
              lambda=NULL) # box-cox transformation, lambda=0 is log-transformatio
summary(fit1)
checkresiduals(fit1)
```


```{r message = FALSE}
fit2 <- Arima(y, order=c(2,1,0),
              seasonal=c(2,0,0),
              include.constant = T,
              lambda=NULL) # box-cox transformation, lambda=0 is log-transformation
summary(fit2)
checkresiduals(fit2)
```


```{r message = FALSE}
fit3 <- Arima(y, order=c(2,1,0),
              xreg=seasonalvars,
              include.constant = T,
              lambda=NULL) # box-cox transformation, lambda=0 is log-transformation
summary(fit3)
checkresiduals(fit3)
```

```{r message = FALSE}
fit4 <- Arima(y, order=c(2,1,0),
              seasonal=c(2,0,0),
              xreg=seasonalvars,
              include.constant = T,
              lambda=NULL) # box-cox transformation, lambda=0 is log-transformation
summary(fit4)
checkresiduals(fit4)
```

In conclusion, we pick ARIMA(2,1,0) with seasonal dummies as our chosen model since it has lowest BIC compared to other.

```{r message = FALSE}
fit_chosen <- Arima(y, order=c(2,1,0),
              xreg=seasonalvars,
              include.constant = T,
              lambda=NULL) # box-cox transformation, lambda=0 is log-transformation
summary(fit_chosen)
```


## 3. Forecasting 2013- 2017 1 step ahead in sample

We test this model on a somewhat validation set to see its performance. Note that this is overfitting since the chosen model is built on this data.

```{r message=FALSE}
no_fcst_mths <- length(window(y, start = c(2013,1), end = c(2017,12)))
no_remain_months <- length(y) - no_fcst_mths
fcst1step <-ts(matrix(rep(NA,no_fcst_mths*3),ncol=3), start=c(2013,1), frequency=12) # to store forecasts
colnames(fcst1step) <- c("mean", "lower", "upper")

for (i in 1:no_fcst_mths){

  temp_mdl <- Arima(y[1:(no_remain_months+i-1)], 
                order = c(2,1,0),
                xreg = seasonalvars[1:(no_remain_months+i-1),],
                include.constant = T,
                biasadj = T,
                lambda = NULL)


  temp <- forecast(temp_mdl, h=1, xreg=matrix(seasonalvars[(no_remain_months+i):(no_remain_months+i),],nrow = 1))
  fcst1step[i,]<-cbind(temp$mean, temp$lower[,"80%"], temp$upper[,"80%"])

}

ts.plot <- ts.union(Actual=window(y, start = c(2010,1)), fcst1step)
autoplot(ts.plot) +
  scale_color_manual(values=rep("black", 4)) +
  ylab("") + xlab("") +
  aes(linetype=series) +
  scale_size_manual(values = c(0.5, 2, 0.75)) +
  scale_linetype_manual(values=c("solid", "dashed", rep("dotted",2))) +
  ts_thm + theme(legend.position="none") + ggtitle("Forecast 1-step ahead for 2013-2017")

```

Evaluate R-squared and RMSE on this period

```{r message=FALSE}
f1err <- ts.union(Actual=window(y,start=c(2013,1)),Fcst=fcst1step[,1])
sse <- sum((f1err[,"Actual"]-f1err[,"Fcst"])^2)
sst <- sum((f1err[,"Actual"]-mean(f1err[,"Actual"]))^2)
OOSR2 <- 1-sse/sst
print(paste0("Out-of-sample RMSE is ",as.character(round(sqrt(sse/length(f1err)),2))))
print(paste0("Out-of-sample R-sqr is ",as.character(round(OOSR2,2))))
```
It is not surprising that the forecast is quite good.

## 4. Forecasting 2018 out of sample  

Forecast 12 period ahead
```{r message = FALSE}
fcst <- forecast(fit_chosen, h=12, xreg = seasonalvars) # forecast 1 to 28 steps (1 years x 12 months) ahead
actual_data <- read_csv(paste(dropbox_path,"\\ECON233\\","dx_2018.csv",sep=""), col_names = FALSE)
colnames(actual_data) <- c("Date", "DX")
actual_data$DX <- log(actual_data$DX)
dx_actual_ts <- ts(actual_data$DX, start = c(1976,1), frequency = 12)
dx_os_ts <- window(dx_actual_ts, start = c(2018, 1), end = c(2018,12), frequecy = 12)

autoplot(ts.union(window(dx_actual_ts, start = c(2015,1), end = c(2018,12)),
                  window(fcst$mean, start = c(2015,1), end = c(2018,12)),
                  window(fcst$lower, start = c(2015,1), end = c(2018,12)),
                  window(fcst$upper, start = c(2015,1), end = c(2018,12)))) +
  scale_color_manual(values=rep("black", 6)) +
  ylab("") + xlab("") +
  aes(linetype=series) +
  scale_size_manual(values = c(0.5, 2, 0.75)) +
  scale_linetype_manual(values=c("solid", "dashed", rep("dotted",4))) +
  ts_thm + theme(legend.position="none") + ggtitle("Forecast 12-period ahead for 2018")

```

Evaluate forecast on real data for 2018

```{r message=FALSE}
f2err <- ts.union(Actual=window(dx_os_ts,start=c(2018,1)),
                  Fcst=window(fcst$mean, start = c(2018,1), end = c(2018,12)))
sse <- sum((f2err[,"Actual"]-f2err[,"Fcst"])^2)
sst <- sum((f2err[,"Actual"]-mean(f2err[,"Actual"]))^2)
OOSR2 <- 1-sse/sst
print(paste0("Out-of-sample RMSE is ",as.character(round(sqrt(sse/12),2))))
print(paste0("Out-of-sample R-sqr is ",as.character(round(OOSR2,2))))
```

Forecast 1-step ahead for the model during 2018.

```{r message=FALSE}
seasonalvars2 <- seasonaldummy(dx_actual_ts)
no_fcst_mths <- length(window(dx_actual_ts, start = c(2018,1), end = c(2018,12)))
no_remain_months <- length(dx_actual_ts) - no_fcst_mths # remove 2 months of jan and feb 2019
fcst1step_3 <-ts(matrix(rep(NA,no_fcst_mths*3),ncol=3), start=c(2018,1), frequency=12) # to store forecasts
colnames(fcst1step_3) <- c("mean", "lower", "upper")

for (i in 1:no_fcst_mths){
  
  temp_mdl <- Arima(dx_actual_ts[1:(no_remain_months+i-1)], 
                     order = c(2,1,0),
                     xreg = seasonalvars2[1:(no_remain_months+i-1),],
                     include.constant = T,
                     biasadj = T,
                     lambda = NULL)
  
  
  
  temp <- forecast(temp_mdl, h=1, xreg=matrix(seasonalvars2[(no_remain_months+i):(no_remain_months+i),],nrow = 1),
                   biasadj = T, lambda = NULL)
  fcst1step_3[i,]<-cbind(temp$mean, temp$lower[,"80%"], temp$upper[,"80%"])
  
}

ts.plot <- ts.union(Actual=window(dx_actual_ts, start = c(2015,1)), fcst1step_3)
autoplot(ts.plot) +
  scale_color_manual(values=rep("black", 4)) +
  ylab("") + xlab("") +
  aes(linetype=series) +
  scale_size_manual(values = c(0.5, 2, 0.75)) +
  scale_linetype_manual(values=c("solid", "dashed", rep("dotted",2))) +
  ts_thm + theme(legend.position="none") + ggtitle("Forecast 1-step ahead for 2018")

```

Evaluate the accuracy and R-squared for this period
```{r message=FALSE}
f1err_2 <- ts.union(Actual=window(dx_actual_ts,start=c(2018,1)),Fcst=fcst1step_3[,1])
sse <- sum((f1err_2[,"Actual"]-f1err_2[,"Fcst"])^2)
sst <- sum((f1err_2[,"Actual"]-mean(f1err_2[,"Actual"]))^2)
OOSR2 <- 1-sse/sst
print(paste0("Out-of-sample RMSE is ",as.character(round(sqrt(sse/length(f1err_2)),2))))
print(paste0("Out-of-sample R-sqr is ",as.character(round(OOSR2,2))))
```